---
title: "Predicting WTI crude oil prices"
subtitle: 'Empirical Project - Causal Inference'
author: "Mads Emil Marker Jungersen & Jeppe Dalgaard Rode"
date: "01/09-2020"
header-includes:
   - \usepackage{amsmath}
   - \usepackage{ textcomp }
   - \usepackage{graphicx}
output:
  pdf_document: default
  html_document: default
  word_document: default
documentclass: article
classoption: a4paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(cowplot)
library(wooldridge)
library(mfx)
library(AER)
library(car)
library(plm)
library(foreign)
library(rddtools)
library(Matching)
library(rgenoud)
library(readxl)
library(tseries)
library(forecast)
library(RSNNS)
library(glmnet)
```

## Importing and fixing data

```{r}
rm(list=ls())
data <- read_excel("Data/Monthly_WTI.xlsx")

##Adding oil supply
supply <- read.csv("Data/Crude_oil_including_lease_condensate_production_World_Monthly.csv", skip = 4, header = T)
supply$Date <- parse_date(paste(substr(supply$Month, 1, 3), '15',substr(supply$Month, 5, 8)),format = "%b %d %Y", locale=locale("en"))
supply <- subset(supply, select = c(Date,Series.ID..INTL.57.1.WORL.TBPD.M.thousand.barrels.per.day))
colnames(supply) <- c("Date", "Supply")
data <- inner_join(data, supply, by="Date")

##Adding kilian index (oil demand)
kilian <- read_excel("Data/kilian.xlsx")
kilian$Date2 <- as.Date(kilian$Date) + 14
kilian <- subset(kilian, select = c(Date2, index))
colnames(kilian) <- c("Date", "KI")
data <- inner_join(data, kilian, by="Date")


##Adding SP500 closing price
SP500 <- read.csv("Data/SP500_closing.csv", header = T)
SP500 <- subset(SP500, select=c(Date, Adj.Close))
SP500$Date2 <- as.Date(as.Date(SP500$Date) + 14)
SP500 <- subset(SP500, select=c(Date2, Adj.Close))
colnames(SP500) <- c("Date", "SP500")
data <- inner_join(data, SP500, by="Date")

#Adding t for trend
data$t <-  seq.int(nrow(data))
data$tsq <- data$t^2
data$tcu <- data$t^3

# Adding a one day lag of oil price
data$Price_1 <- c(NA,data$Price[-length(data$Price)])

#####Adding Geopolitical Events
#Iraq invades kuwait (one in aug-sep 1990)
data$IIK <- ifelse(data$Date >= as.Date("1990-08-01") & data$Date < as.Date("1990-10-01"),1,0)
#Asian financial crisis (one in july 1997-dec 1998)
data$AFC <- ifelse(data$Date >= as.Date("1997-07-01") & data$Date < as.Date("1999-01-01"),1,0)
#9/11 attacks (one in sep-nov 2001)
data$NE <- ifelse(data$Date >= as.Date("2001-09-01") & data$Date < as.Date("2001-12-01"),1,0)
#Global financial collapse (One in sep 2008 to feb 2009)
data$GFC <- ifelse(data$Date >= as.Date("2008-09-01") & data$Date < as.Date("2009-03-01"),1,0)
#Global pandamic crisis (One in 2020)
data$GPC <- ifelse(data$Date >= as.Date("2020-01-01"),1,0)


#####Adding Dummy-variables for month
data$feb <- ifelse(grepl("-02",data$Date),1,0)
data$mar <- ifelse(grepl("-03",data$Date),1,0)
data$apr <- ifelse(grepl("-04",data$Date),1,0)
data$may <- ifelse(grepl("-05",data$Date),1,0)
data$jun <- ifelse(grepl("-06",data$Date),1,0)
data$jul <- ifelse(grepl("-07",data$Date),1,0)
data$aug <- ifelse(grepl("-08",data$Date),1,0)
data$sep <- ifelse(grepl("-09",data$Date),1,0)
data$oct <- ifelse(grepl("-10",data$Date),1,0)
data$nov <- ifelse(grepl("-11",data$Date),1,0)
data$dec <- ifelse(grepl("-12",data$Date),1,0)

########PLOT THE MONTHLY DATA
#jpeg(file="monthly.jpeg")
plot(ts(data$Price,start = c(1986,1), frequency = 12), ylab="WTI Crude Oil Price", main = "Monthly WTI crude oil spot price")
#dev.off()

```

## Augmented Dickey Fuller test for unit-root auto $k$

```{r}
lag_orders <- numeric(4)
test_statistics <- numeric(4)
p_vals <- numeric(4)
for (i in 2:5){
  lag_orders[i-1] <- adf.test(ts(data[i]))$parameter
  test_statistics[i-1] <- adf.test(ts(data[i]))$statistic
  p_vals[i-1] <- adf.test(ts(data[i]))$p.value
}

Table <- cbind(lag_orders,test_statistics,p_vals)
rownames(Table)<- c("Price", "Supply", "KI", "SP500")
colnames(Table) <- c("# Lags", "Test Statistic", "P-value")
Table %>%
  kable(digits = 3, "latex", booktabs = T)%>%
  kable_styling()%>%
  add_header_above(c("Augmented Dickey Fuller test"=4))

```
## Augmented Dickey Fuller test for unit-root auto $k=0$

```{r}
lag_orders <- numeric(4)
test_statistics <- numeric(4)
p_vals <- numeric(4)
for (i in 2:5){
  lag_orders[i-1] <- adf.test(ts(data[i]),k=0)$parameter
  test_statistics[i-1] <- adf.test(ts(data[i]),k=0)$statistic
  p_vals[i-1] <- adf.test(ts(data[i]),k=0)$p.value
}

Table <- cbind(lag_orders,test_statistics,p_vals)
rownames(Table)<- c("Price", "Supply", "KI", "SP500")
colnames(Table) <- c("# Lags", "Test Statistic", "P-value")
Table %>%
  kable(digits = 3, "latex", booktabs = T)%>%
  kable_styling()%>%
  add_header_above(c("Augmented Dickey Fuller test"=4))

```


## First difference data that have unit root

```{r}
data$D_Price <- c(NA,diff(data$Price))
data$D_Supply <- c(NA,diff(data$Supply))
data$D_KI <- c(NA,diff(data$KI))
data$D_SP500 <- c(NA,diff(data$SP500))
```


## Lagging variables once

```{r}
data$Price_1 <-  c(NA,data$Price[-length(data$Price)])
data$Supply_1 <-  c(NA,data$Supply[-length(data$Supply)])
data$KI_1 <-  c(NA,data$KI[-length(data$KI)])
data$SP500_1 <-  c(NA,data$SP500[-length(data$SP500)])
data$D_Price_1 <-  c(NA,data$D_Price[-length(data$D_Price)])
data$D_Supply_1 <-  c(NA,data$D_Supply[-length(data$D_Supply)])
data$D_KI_1 <-  c(NA,data$D_KI[-length(data$D_KI)])
data$D_SP500_1 <-  c(NA,data$D_SP500[-length(data$D_SP500)])
```
## Lagging variables twice

```{r}
data$Price_2 <-  c(NA,data$Price_1[-length(data$Price_1)])
data$D_Price_2 <-  c(NA,data$D_Price_1[-length(data$D_Price_1)])
data$D_Supply_2 <-  c(NA,data$D_Supply_1[-length(data$D_Supply_1)])
data$D_KI_2 <-  c(NA,data$D_KI_1[-length(data$D_KI_1)])
data$D_SP500_2 <-  c(NA,data$D_SP500_1[-length(data$D_SP500_1)])
```

## Adding more lags 
```{r}
##PRICE
data$D_Price_3 <-  c(NA,data$D_Price_2[-length(data$D_Price_2)])
data$D_Price_4 <-  c(NA,data$D_Price_3[-length(data$D_Price_3)])
data$D_Price_5 <-  c(NA,data$D_Price_4[-length(data$D_Price_4)])
data$D_Price_6 <-  c(NA,data$D_Price_5[-length(data$D_Price_5)])

##SUPPLY
data$D_Supply_3 <-  c(NA,data$D_Supply_2[-length(data$D_Supply_2)])
data$D_Supply_4 <-  c(NA,data$D_Supply_3[-length(data$D_Supply_3)])
data$D_Supply_5 <-  c(NA,data$D_Supply_4[-length(data$D_Supply_4)])
data$D_Supply_6 <-  c(NA,data$D_Supply_5[-length(data$D_Supply_5)])

## KI
data$D_KI_3 <-  c(NA,data$D_KI_2[-length(data$D_KI_2)])
data$D_KI_4 <-  c(NA,data$D_KI_3[-length(data$D_KI_3)])
data$D_KI_5 <-  c(NA,data$D_KI_4[-length(data$D_KI_4)])
data$D_KI_6 <-  c(NA,data$D_KI_5[-length(data$D_KI_5)])

## SP500
data$D_SP500_3 <-  c(NA,data$D_SP500_2[-length(data$D_SP500_2)])
data$D_SP500_4 <-  c(NA,data$D_SP500_3[-length(data$D_SP500_3)])
data$D_SP500_5 <-  c(NA,data$D_SP500_4[-length(data$D_SP500_4)])
data$D_SP500_6 <-  c(NA,data$D_SP500_5[-length(data$D_SP500_5)])

```



## Exporting data

```{r}
save(data, file = "data.RData")
```



## Forcasting models

Making out-of-sample where the data from January 1986 uptil and including December 2013 are used as training the model, and the data from January 2014 up til and including August 2020 are used for testing the model. This roughly corresponds to a $80/20\%$ split of the data. 

```{r}
train <- subset(data, Date < as.Date("2014-01-01"))
test <- subset(data, Date >= as.Date("2014-01-01"))
```

First want to determine the number of lags we should use of the dependent variable

```{r}
LM <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + IIK + AFC + NE + GFC, data=subset(train,t>7))
summary(LM)
LM1 <- lm(D_Price ~ D_Price_1 + IIK + AFC + NE + GFC, data=subset(train,t>7))
anova(LM1,LM)
```

We will be using $6$ lags, and compare it with the $AR(1)$ model.

### (i) Testing for the explanatory variables when using 6 lags.

#### Does supply granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_Supply_1 +D_Supply_2 + D_Supply_3 + D_Supply_4 + D_Supply_5 + D_Supply_6, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6, data=subset(train,t>7))
anova(LMtil,LMUD)
```
Dosent seem like it.

#### Does KI granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_KI_1 + D_KI_2 + D_KI_3 + D_KI_4 + D_KI_5 + D_KI_6, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6, data=subset(train,t>7))
anova(LMtil,LMUD)
```
Dosent seem like it - however close to being significant.

#### Does SP500 granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + D_SP500_2 + D_SP500_3 + D_SP500_4 + D_SP500_5 + D_SP500_6, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6, data=subset(train,t>7))
anova(LMtil,LMUD)
```
YES!

```{r}
#HOW MANY LAGS?
#1?
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + D_SP500_2 + D_SP500_3 + D_SP500_4 + D_SP500_5 + D_SP500_6, data=subset(train,t>7))
#summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6+ D_SP500_1, data=subset(train,t>7))
anova(LMtil,LMUD)
# One is enough
```

#### Does Supply granger cause the price when conditioning on SP500?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + D_Supply_1 + D_Supply_2 + D_Supply_3 + D_Supply_4 + D_Supply_5 + D_Supply_6, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1, data=subset(train,t>7))
anova(LMtil,LMUD)
```
No.


#### Does KI granger cause the price when conditioning on SP500?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + D_KI_1 + D_KI_2 + D_KI_3 + D_KI_4 + D_KI_5 + D_KI_6, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1, data=subset(train,t>7))
anova(LMtil,LMUD)
```
No.






### (ii) Testing for the explanatory variables when using 6 lags and dummies.

#### Does supply granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_Supply_1 +D_Supply_2 + D_Supply_3 + D_Supply_4 + D_Supply_5 + D_Supply_6+ IIK + AFC + NE + GFC, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6+ IIK + AFC + NE + GFC, data=subset(train,t>7))
anova(LMtil,LMUD)
```
Dosent seem like it.

#### Does KI granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_KI_1 + D_KI_2 + D_KI_3 + D_KI_4 + D_KI_5 + D_KI_6+ IIK + AFC + NE + GFC, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6+ IIK + AFC + NE + GFC, data=subset(train,t>7))
anova(LMtil,LMUD)
```
Dosent seem like it - however close to being significant.

#### Does SP500 granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + D_SP500_2 + D_SP500_3 + D_SP500_4 + D_SP500_5 + D_SP500_6 + IIK + AFC + NE + GFC, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6+ IIK + AFC + NE + GFC, data=subset(train,t>7))
anova(LMtil,LMUD)
```
SP500_t-1 does.

```{r}
#HOW MANY LAGS?
#1?
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + D_SP500_2 + D_SP500_3 + D_SP500_4 + D_SP500_5 + D_SP500_6, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6+ D_SP500_1, data=subset(train,t>7))
anova(LMtil,LMUD)
# One is enough
```

#### Does Supply granger cause the price when conditioning on SP500?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + D_Supply_1 + D_Supply_2 + D_Supply_3 + D_Supply_4 + D_Supply_5 + D_Supply_6 + IIK + AFC + NE + GFC, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + IIK + AFC + NE + GFC, data=subset(train,t>7))
anova(LMtil,LMUD)
```
No.


#### Does KI granger cause the price when conditioning on SP500?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1 + D_KI_1 + D_KI_2 + D_KI_3 + D_KI_4 + D_KI_5 + D_KI_6 + IIK + AFC + NE + GFC, data=subset(train,t>7))
summary(LMUD)
LMtil <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC, data=subset(train,t>7))
anova(LMtil,LMUD)
```
No.



### (iii) Testing for the explanatory variables when using 1 lag.

#### Does supply granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Supply_1, data=train)
summary(LMUD)
```

Dosent seem like it.

#### Does KI granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_KI_1, data=train)
summary(LMUD)
```
Dosent seem like it - however close to being significant.

#### Does SP500 granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_SP500_1, data=train)
summary(LMUD)
```
YES!


#### Does Supply granger cause the price when conditioning on SP500?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 +D_SP500_1 + D_Supply_1, data=train)
summary(LMUD)

```
No.

#### Does KI granger cause the price when conditioning on SP500?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 +D_SP500_1 + D_KI_1, data=train)
summary(LMUD)

```
No.




### (iv) Testing for the explanatory variables when using 1 lag and dummies.

#### Does supply granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_Supply_1 + IIK + AFC + NE + GFC, data=train)
summary(LMUD)
```
Dosent seem like it.

#### Does KI granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_KI_1+ IIK + AFC + NE + GFC, data=train)
summary(LMUD)
```
Dosent seem like it - however close to being significant.

#### Does SP500 granger cause price?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
summary(LMUD)
```
YES!


#### Does Supply granger cause the price when conditioning on SP500?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 +D_SP500_1 + D_Supply_1, data=train)
summary(LMUD)

```
No.

#### Does KI granger cause the price when conditioning on SP500?

```{r}
LMUD <- lm(D_Price ~ D_Price_1 +D_SP500_1 + D_KI_1+ IIK + AFC + NE + GFC, data=train)
summary(LMUD)

```
No.




## Defining the models

### $AR(6)+X$

```{r}
AR6X <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1, data=train)
summary(AR6X)

## Making predictions
predictions <- predict(AR6X, test)
predictions <- predictions + test$Price_1

## Error calculating
RMSE_AR6X <- sqrt(mean((test$Price - predictions)^2))
MAE_AR6X <- mean(abs(test$Price - predictions))

```

### $AR(6)+X$ and dummies

```{r}
AR6Xd <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
summary(AR6Xd)

## Making predictions
predictions <- predict(AR6Xd, test)
predictions <- predictions + test$Price_1

## Error calculating
RMSE_AR6Xd <- sqrt(mean((test$Price - predictions)^2))
MAE_AR6Xd <- mean(abs(test$Price - predictions))

```



### $AR(1)+X$

```{r}
AR1X <- lm(D_Price ~ D_Price_1 + D_SP500_1, data=train)
summary(AR1X)

## Making predictions
predictions <- predict(AR1X, test)
predictions <- predictions + test$Price_1

## Error calculating
RMSE_AR1X <- sqrt(mean((test$Price - predictions)^2))
MAE_AR1X <- mean(abs(test$Price - predictions))
```


### $AR(1)+X$ and dummies

```{r}
AR1Xd <- lm(D_Price ~ D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
summary(AR1Xd)

## Making predictions
predictions <- predict(AR1Xd, test)
predictions <- predictions + test$Price_1

## Error calculating
RMSE_AR1Xd <- sqrt(mean((test$Price - predictions)^2))
MAE_AR1Xd <- mean(abs(test$Price - predictions))

```



### $AR(6)$

```{r}
AR6 <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6, data=train)
summary(AR6)

## Making predictions
predictions <- predict(AR6, test)
predictions <- predictions + test$Price_1

## Error calculating
RMSE_AR6 <- sqrt(mean((test$Price - predictions)^2))
MAE_AR6 <- mean(abs(test$Price - predictions))

```

### $AR(1)$

```{r}
AR1 <- lm(D_Price ~ D_Price_1, data=train)
summary(AR1)

## Making predictions
predictions <- predict(AR1, test)
predictions <- predictions + test$Price_1

## Error calculating
RMSE_AR1 <- sqrt(mean((test$Price - predictions)^2))
MAE_AR1 <- mean(abs(test$Price - predictions))

```

### Random Walk

```{r}
## Making predictions
predictions <- test$Price_1

## Error calculating
RMSE_RW <- sqrt(mean((test$Price - predictions)^2))
MAE_RW <- mean(abs(test$Price - predictions))
```

#### Creating a table summarizing errors

```{r}
RMSE <- cbind(RMSE_AR6X,RMSE_AR1X,RMSE_AR6Xd,RMSE_AR1Xd,RMSE_AR6,RMSE_AR1,RMSE_RW)
MAE <- cbind(MAE_AR6X,MAE_AR1X,MAE_AR6Xd,MAE_AR1Xd,MAE_AR6,MAE_AR1,MAE_RW)
Table <- rbind(RMSE,MAE)
colnames(Table) <- c("AR6X", "AR1X", "AR6Xd", "AR1Xd", "AR(6)", "AR(1)", "Random Walk")
rownames(Table) <- c("RMSE", "MAE")
Table %>%
  kable(digits = 3, "latex", booktabs = T)%>%
  kable_styling(latex_options="scale_down")%>%
  add_header_above(c("Forecassting performance"=8))
```

### Tryng the same models, but leaving out the data for $2020$ (pandemic)

```{r}
testCORONA <- subset(data, Date >= as.Date("2014-01-01") & Date < as.Date("2020-01-01"))
```

### $AR(6)+X$ CORONA

```{r}
AR6X <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1, data=train)
#summary(AR6X)

## Making predictions
predictions <- predict(AR6X, testCORONA)
predictions <- predictions + testCORONA$Price_1

## Error calculating
RMSE_AR6X <- sqrt(mean((testCORONA$Price - predictions)^2))
MAE_AR6X <- mean(abs(testCORONA$Price - predictions))

```

### $AR(1)+X$ CORONA

```{r}
AR1X <- lm(D_Price ~ D_Price_1 + D_SP500_1, data=train)
#summary(AR1X)

## Making predictions
predictions <- predict(AR1X, testCORONA)
predictions <- predictions + testCORONA$Price_1

## Error calculating
RMSE_AR1X <- sqrt(mean((testCORONA$Price - predictions)^2))
MAE_AR1X <- mean(abs(testCORONA$Price - predictions))

```


### $AR(6)+X$ and dummies CORONA

```{r}
AR6Xd <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
summary(AR6Xd)

## Making predictions
predictions <- predict(AR6Xd, testCORONA)
predictions <- predictions + testCORONA$Price_1

## Error calculating
RMSE_AR6Xd <- sqrt(mean((testCORONA$Price - predictions)^2))
MAE_AR6Xd <- mean(abs(testCORONA$Price - predictions))

```

### $AR(1)+X$ and dummies CORONA

```{r}
AR1Xd <- lm(D_Price ~ D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
summary(AR1Xd)

## Making predictions
predictions <- predict(AR1Xd, testCORONA)
predictions <- predictions + testCORONA$Price_1

## Error calculating
RMSE_AR1Xd <- sqrt(mean((testCORONA$Price - predictions)^2))
MAE_AR1Xd <- mean(abs(testCORONA$Price - predictions))

```


### $AR(6)$ CORONA

```{r}
AR6 <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6, data=train)
#summary(AR6)

## Making predictions
predictions <- predict(AR6, testCORONA)
predictions <- predictions + testCORONA$Price_1

## Error calculating
RMSE_AR6 <- sqrt(mean((testCORONA$Price - predictions)^2))
MAE_AR6 <- mean(abs(testCORONA$Price - predictions))

```

### $AR(1)$ CORONA

```{r}
AR1 <- lm(D_Price ~ D_Price_1, data=train)
#summary(AR1)

## Making predictions
predictions <- predict(AR1, testCORONA)
predictions <- predictions + testCORONA$Price_1

## Error calculating
RMSE_AR1 <- sqrt(mean((testCORONA$Price - predictions)^2))
MAE_AR1 <- mean(abs(testCORONA$Price - predictions))

```

### Random Walk CORONA

```{r}
## Making predictions
predictions <- testCORONA$Price_1

## Error calculating
RMSE_RW <- sqrt(mean((testCORONA$Price - predictions)^2))
MAE_RW <- mean(abs(testCORONA$Price - predictions))
```

#### Creating a table summarizing errors

```{r}
RMSE <- cbind(RMSE_AR6X,RMSE_AR1X,RMSE_AR6Xd,RMSE_AR1Xd,RMSE_AR6,RMSE_AR1,RMSE_RW)
MAE <- cbind(MAE_AR6X,MAE_AR1X,MAE_AR6Xd,MAE_AR1Xd,MAE_AR6,MAE_AR1,MAE_RW)
Table <- rbind(RMSE,MAE)
colnames(Table) <- c("AR6X", "AR1X", "AR6Xd", "AR1Xd", "AR(6)", "AR(1)", "Random Walk")
rownames(Table) <- c("RMSE", "MAE")
Table %>%
  kable(digits = 3, "latex", booktabs = T)%>%
  kable_styling(latex_options="scale_down")%>%
  add_header_above(c("Forecassting performance leaving out 2020"=8))
```

## Testing for serial correlation and heteroskedasticity

### AR6X

```{r}
AR6X <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1, data=train)
u <- AR6X$residuals
u <- c(NA,NA,NA,NA,NA,NA,NA,u)
u_1 <- c(NA,u[-length(u)])
res <- summary(lm(u~u_1 + D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1, data=train))
BG_pval_AR6X <- res$coefficients[2,4]
DW_pval_AR6X <- dwtest(AR6X)$p.val
u2 <- u^2
BP <- summary(lm(u2~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1, data=train))
BP_pval_AR6X <- 0.0026
```



### AR1X

```{r}
AR1X <- lm(D_Price ~ D_Price_1 + D_SP500_1, data=train)
u <- AR1X$residuals
u <- c(NA,NA,u)
u_1 <- c(NA,u[-length(u)])
res <- summary(lm(u~u_1 + D_Price_1 + D_SP500_1, data=train))
BG_pval_AR1X <- res$coefficients[2,4]
DW_pval_AR1X <- dwtest(AR1X)$p.val
u2 <- u^2
BP <- summary(lm(u2~ D_Price_1 + D_SP500_1, data=train))
BP_pval_AR1X <- 0.005
```


### AR6Xd

```{r}
AR6Xd <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
u <- AR6Xd$residuals
u <- c(NA,NA,NA,NA,NA,NA,NA,u)
u_1 <- c(NA,u[-length(u)])
res <- summary(lm(u~u_1 + D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC, data=train))
BG_pval_AR6Xd <- res$coefficients[2,4]
DW_pval_AR6Xd <- dwtest(AR6Xd)$p.val
u2 <- u^2
BP <- summary(lm(u2~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC, data=train))
BP_pval_AR6Xd <- 0.0001
```
### AR1Xd

```{r}
AR1Xd <- lm(D_Price ~ D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
u <- AR1Xd$residuals
u <- c(NA,NA,u)
u_1 <- c(NA,u[-length(u)])
res <- summary(lm(u~u_1 + D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC, data=train))
BG_pval_AR1Xd <- res$coefficients[2,4]
DW_pval_AR1Xd <- dwtest(AR1Xd)$p.val
u2 <- u^2
BP <- summary(lm(u2~ D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC, data=train))
BP_pval_AR1Xd <- 0.0001
```
### AR6

```{r}
AR6 <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6, data=train)
u <- AR6$residuals
u <- c(NA,NA,NA,NA,NA,NA,NA,u)
u_1 <- c(NA,u[-length(u)])
res <- summary(lm(u~u_1 + D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6, data=train))
BG_pval_AR6 <- res$coefficients[2,4]
DW_pval_AR6 <- dwtest(AR6)$p.val
u2 <- u^2
BP <- summary(lm(u2~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6, data=train))
BP_pval_AR6 <- 0.0002
```
### AR1

```{r}
AR1 <- lm(D_Price ~ D_Price_1, data=train)
u <- AR1$residuals
u <- c(NA,NA,u)
u_1 <- c(NA,u[-length(u)])
res <- summary(lm(u~u_1 + D_Price_1, data=train))
BG_pval_AR1 <- res$coefficients[2,4]
DW_pval_AR1 <- dwtest(AR1)$p.val
u2 <- u^2
BP <- summary(lm(u2~ D_Price_1, data=train))
BP_pval_AR1 <- 0.0004
```


#### Table summarizing results from tests
```{r}
BG <- cbind(BG_pval_AR6X,BG_pval_AR1X,BG_pval_AR6Xd,BG_pval_AR1Xd,BG_pval_AR6,BG_pval_AR1)
DW <- cbind(DW_pval_AR6X,DW_pval_AR1X,DW_pval_AR6Xd,DW_pval_AR1Xd,DW_pval_AR6,DW_pval_AR1)
BP <- cbind(BP_pval_AR6X,BP_pval_AR1X,BP_pval_AR6Xd,BP_pval_AR1Xd,BP_pval_AR6,BP_pval_AR1)
Table <- rbind(BG,DW,BP)
colnames(Table) <- c("AR6X", "AR1X", "AR6Xd", "AR1Xd", "AR(6)", "AR(1)")
rownames(Table) <- c("BG", "DW", "BP")
Table %>%
  kable(digits = 3, "latex", booktabs = T)%>%
  kable_styling()%>%
  add_header_above(c("Tests for serial sorrelation and heteroskedasticity (p-vals)"=7))
```

## Things in the discussion

```{r}
plot(ts(test$Price,start = c(2014,1), frequency = 12), ylab="WTI Crude Oil Price", main = "Monthly WTI crude oil spot price")
plot(ts(test$SP500,start = c(2014,1), frequency = 12), ylab="WTI Crude Oil Price", main = "Monthly WTI crude oil spot price")
```

### Seasonality and trend

```{r}
# First adding a linear trend
AR6Xd <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC + t, data=train)
AR1Xd <- lm(D_Price ~ D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC +t, data=train)
AR6X <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+t, data=train)
AR1X <- lm(D_Price ~ D_Price_1 + D_SP500_1+t, data=train)
coeftest(AR6Xd, vcov=vcovHAC(AR6Xd))
coeftest(AR1Xd, vcov=vcovHAC(AR1Xd))
coeftest(AR6X, vcov=vcovHAC(AR6X))
coeftest(AR1X, vcov=vcovHAC(AR1X))
# None are significant at a 0.01 significance level

# Seasonality
AR6Xd <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
AR6Xd_s <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+ IIK + AFC + NE + GFC + feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec, data=train)
waldtest(AR6Xd,AR6Xd_s,vcov = vcovHAC(AR6Xd_s))
#NOT SIGNIFICANT
AR1Xd <- lm(D_Price ~ D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC, data=train)
AR1Xd_s <- lm(D_Price ~ D_Price_1 + D_SP500_1+ IIK + AFC + NE + GFC +feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec, data=train)
waldtest(AR1Xd,AR1Xd_s,vcov = vcovHAC(AR1Xd_s))
#NOT SIGNIFICANT
AR6X <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1, data=train)
AR6X_s <- lm(D_Price ~ D_Price_1 + D_Price_2 + D_Price_3 + D_Price_4 + D_Price_5 + D_Price_6 + D_SP500_1+feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec, data=train)
waldtest(AR6X,AR6X_s,vcov = vcovHAC(AR6X_s))
#NOT SIGNIFICANT
AR1X <- lm(D_Price ~ D_Price_1 + D_SP500_1, data=train)
AR1X_s <- lm(D_Price ~ D_Price_1 + D_SP500_1+feb+mar+apr+may+jun+jul+aug+sep+oct+nov+dec, data=train)
waldtest(AR1X,AR1X_s,vcov = vcovHAC(AR1X_s))
#Not significant
```



